\documentclass[11pt]{article}
\usepackage{amsmath, amssymb, graphicx, booktabs, hyperref, natbib}
\usepackage[a4paper, margin=1in]{geometry}

\title{Building Local-Language Economic Narrative Indices: \\
The BENI v1 Pipeline from Raw News to Validated Measurement}

\author{Ann Naser Nabil \\
Department of Economics, Jahangirnagar University \\
\texttt{lila.lab0x@gmail.com}}

\date{June 2026}

\begin{document}

\maketitle

\begin{abstract}
Economic narrative indices---monthly measures of economic news salience---are well-established for English-language media but remain scarce for low-resource languages. We present the BENI v1 method pipeline, a reproducible workflow for transforming Bangla-language news into economic narrative measurements for Bangladesh.

The BENI v1 database combines dated Potrika articles from 2014--2020 with post-2020 BNAD articles, producing 1,467,705 harmonised article records with source metadata, deduplication fields, economic seed labels, and model-ready text. The current frozen validation set contains 186 clean reference labels mapped to canonical BENI v1 article IDs; the remaining 114 ambiguous rows are conservatively excluded from clean validation.

In the frozen BENI v1 validation run, TF-IDF + Logistic Regression achieves 88.17\% accuracy, $\kappa = 0.736$, precision of 92.73\%, recall of 73.91\%, and F1 of 0.823 on the Economic class. The corresponding monthly BENI v1 index is built from 1,453,510 deduplicated article predictions and provides both article-weighted and source-balanced specifications.

The contribution is a complete low-resource measurement template: database harmonisation, weak labeling, LLM-assisted reference annotation, frozen validation, article-level prediction, index construction, and macro-validation checks. Legacy Potrika calibration results are retained as prototype evidence only; the final BENI v1 outputs are grounded in the repaired corpus and frozen validation set.

\textbf{Keywords:} economic narratives, Bangla NLP, low-resource text classification, annotation cost, narrative index, BENI

\textbf{JEL Codes:} C53, C55, E37
\end{abstract}

\input{drafts/introduction}

\input{drafts/related_work}

\input{drafts/methods}

\input{drafts/results}

\input{drafts/discussion}

\input{drafts/template}

\input{drafts/conclusion}

\paragraph{Acknowledgments.}
The author thanks the anonymous developers of the Potrika corpus and the BNLP dataset for making Bangla NLP research possible.

\bibliographystyle{plainnat}
\bibliography{references}

\appendix
\section{Annotation Prompt}\label{app:annotation-prompt}

The full annotation prompt used for Claude Sonnet 4 is available in the supplementary materials and the GitHub repository (\url{https://github.com/ann-nabil/economic-narrative-indices}).

\section{Economic Keyword List}\label{app:keywords}

The full list of 32 Bangla economic terms used for keyword-based pseudo-label generation is available in the supplementary materials.

\end{document}
